---
title: "DMBA data cleaning"
author: "Joshua"
date: "5/2/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r cars}
library(tidyverse)
dat<-read.csv("/Users/joonghochoi/Desktop/hackathon_scouting.csv")
date<-dat$game_date
a<-which(substr(date,1,4)=="2019")
b<-which(substr(date,1,4)=="2020")
c<-union(a,b)
```
#Now has filtered 2019 and 2020 data with only relevant columns.
#Now convert categorical to numerical


#Neural Network
```{r}
dat7 <-subset(dat,select= -c(Clock,Player,Player.2,Home.Team.Goals,Away.Team.Goals,Detail.1,Detail.2,Detail.3,Detail.4))
dat8<-dat7[c,]
dat9<-subset(dat8,select= -c(game_date,X.Coordinate.2,Y.Coordinate.2))
unique(dat9$Event)
k<-c("Play","Shot","Goal")
dat10<-dat9%>%filter(Event %in% k)
dat11<-dat10
dat11$Event[dat11$Event=="Play"]<-0
dat11$Event[dat11$Event=="Shot"]<-1
dat11$Event[dat11$Event=="Goal"]<-2
dat12 <- dat11[dat11$Event !=0,]
dat13 <- fastDummies::dummy_cols(dat12)
```

```{r}
dat14<-dat13
dat14$y="meow"
dat14[which(dat14$Event==1),]$y="Shot"
dat14[which(dat14$Event==2),]$y="Goal"
dat14$y<-as.factor(dat14$y)
view(dat14)
```

```{r}
tr_rows = sample(row.names(dat14),dim(dat14)[1]*0.1)
ts_rows = setdiff(row.names(dat14),tr_rows)
hockey_tr = dat14[tr_rows,]
hockey_ts = dat14[ts_rows,]
```

```{r}
library(caret)
ctrl = trainControl(method="cv",      # simple cross-validation
                    number = 10,      # 10 folds
                    classProbs=TRUE  # this should be included so that the 
                    #    algorithm will store predicted probabilities 
                    #     for later generating ROC plot.
)
```

```{r}
set.seed(100)
nnetFit = train( y ~ X.Coordinate + Y.Coordinate + Period + Home.Team.Skaters + Away.Team.Skaters, data=hockey_tr, 
                 method = "nnet",   
                 tuneGrid=expand.grid(decay = c(0, 0.05, 0.1, 1, 2), size = c(1:5, 10)),
                 #tuneLength=5,
                 preProc = c("center","scale"),  # Center and scale data
                 metric ="ROC",
                 trControl = ctrl)
```

```{r}
nnetFit$finalModel
plot(nnetFit)
```

```{r}
testNNET = predict(nnetFit, newdata = hockey_ts)
testNNETProb = predict(nnetFit, newdata= hockey_ts, type = "prob")
trainNNET= predict(nnetFit, newdata = hockey_tr)
```

```{r}
confusionMatrix(data = testNNET, hockey_ts$y)
confusionMatrix(data = testNNET, hockey_ts$y)$overall[1]
confusionMatrix(data = trainNNET,hockey_tr$y)$overall[1]
```
#KNN
```{r}
dat16<-fastDummies::dummy_cols(dat11)
dat17<-dat16
dat17$y="meow"
dat17[which(dat17$Event==0),]$y="Play"
dat17[which(dat17$Event==1),]$y="Shot"
dat17[which(dat17$Event==2),]$y="Goal"
dat17$y<-as.factor(dat17$y)
view(dat17)
```

```{r}
tr2_rows = sample(row.names(dat17),dim(dat17)[1]*0.3)
ts2_rows = setdiff(row.names(dat17),tr2_rows)
hockey_tr2 = dat17[tr2_rows,]
hockey_ts2 = dat17[ts2_rows,]
```

```{r}
set.seed(100)

knn_hockey <- train(y ~ X.Coordinate + Y.Coordinate + Period + Home.Team.Skaters + Away.Team.Skaters, data=hockey_tr2, 
                   method = "knn",
                   preProcess = c("center", "scale"), 
                   tuneGrid=expand.grid(k=1:20),
                   trControl = trainControl(method = "cv", number=5))
```

```{r}
plot(knn_hockey)
knn_hockey
```

```{r}
testknn = predict(knn_hockey, newdata = hockey_ts2)
testknnProb = predict(knn_hockey, newdata= hockey_ts2, type = "prob")
trainknn= predict(knn_hockey, newdata = hockey_tr2)
```

```{r}
confusionMatrix(data = testknn, hockey_ts2$y)
```


#Old SVM

results2<-results1
results2$y="meow"
results2[which(results2$Event==0),]$y="Play"
results2[which(results2$Event==1),]$y="Shot"
results2[which(results2$Event==2),]$y="Goal"
results2$y<-as.factor(results2$y)
results2[results2$y !="Play",]
view(results2)

partition data in training (60%) and test data (40%)
set.seed(100)
tr_rows = sample(row.names(results2),dim(results2)[1]*0.6)
ts_rows = setdiff(row.names(results2),tr_rows)
hockey_tr = results2[tr_rows,]
hockey_ts = results2[ts_rows,]


setting the resampling method
ctrl = trainControl(method="cv",      # simple cross-validation
                    number = 10,      # 10 folds
                    classProbs=TRUE  # this should be included so that the 
                    #    algorithm will store predicted probabilities 
                    #     for later generating ROC plot.
                    )


######################## SVM Linear ##########
training svm model - Linear kernel
set.seed(100)
svmFit = train(y ~ ., data=hockey_tr, 
               method="svmLinear",   # linear kernel
               tuneGrid=expand.grid(C = c(0.01,0.1,1,10,100)),
               preProc=c("center","scale"),  # Center and scale data
               trControl=ctrl)

svmFit$finalModel
plot(svmFit)

testlinearSVM = predict(svmFit, newdata=hockey_ts)
testlinearSVMProb = predict(svmFit, newdata=hockey_ts, type="prob")

confusionMatrix(data=testlinearSVM, hockey_ts$y)

#need to find better summary function. Cannot use twoClassSummary

#ignore codes below.


ctrl = trainControl(method="cv",      # simple cross-validation
                    number = 10,      # 10 folds
                    summaryFunction=twoClassSummary,	# Use AUC to pick the best model
                    classProbs=TRUE  # this should be included so that the 
                    #    algorithm will store predicted probabilities 
                    #     for later generating ROC plot.
                    )
                    


results2$y<-as.factor(results2$y)

results2[which(results2$Event==0),]$y="Play"

results2[which(results2$Event==0),]$y=as.factor("Play")
results2[which(results2$Event==1),]$y=as.factor("Shot")
results2[which(results2$Event==2),]$y=as.factor("Goal")
results2<-results2[complete.cases(results2),]


