---
output:
  pdf_document: default
  html_document: default
---
Step 0
```{r}
library("caret")
library("pROC")
library("kernlab")
credit_data <- read.csv("credit_data.csv", stringsAsFactors = TRUE)
credit_data$credit_history <- ordered(credit_data$credit_history,
                                      levels = c("critical", "poor", "good","very good","perfect"))
credit_data$employment_duration <- 
  ordered(credit_data$employment_duration,
  levels = c("unemployed" , "< 1 year", "1 - 4 years", "4 - 7 years","> 7 years"))

```
Step 1

KNN
```{r}
set.seed(100)

knn_credit <- train(default ~ ., data=credit_data, 
                   method = "knn",
                   preProcess = c("center", "scale"), 
                   tuneGrid=expand.grid(k=1:50),
                   trControl = trainControl(method = "cv", number=10))

pred_knn <- predict(knn_credit, newdata = credit_data)
confusionMatrix(data = pred_knn, credit_data$default)
```
SVM linear
```{r}
set.seed(23)

svm_credit = train(default ~ ., data=credit_data, 
               method="svmLinear",
               tuneGrid=expand.grid(C = c(0.01,0.1,1,10)),
               preProcess=c("center","scale"),
               trControl= trainControl(method = "cv", number=10, classProbs = TRUE))

pred_svm <-predict(svm_credit, newdata = credit_data)
confusionMatrix(data = pred_svm, credit_data$default)
```
Decision tree
```{r}
set.seed(100)

C5.0_credit <- train(default ~ ., data=credit_data,
                     method = "C5.0",
                     preProcess = c("center", "scale"),
                     tuneGrid = expand.grid( .winnow = c(TRUE,FALSE), .trials=1, .model="tree"),
                     trControl = trainControl(method = "cv", number=10))

pred_tree <- predict(C5.0_credit, newdata = credit_data)
confusionMatrix(data = pred_tree, credit_data$default)
```
Step 2

KNN
```{r}
pred_probs_knn <-predict(knn_credit, credit_data, type = "prob")
knnROC <- roc(credit_data$default, 
              pred_probs_knn[,"yes"],
              levels = c('no','yes'),
              direction ='<')
```
SVM linear
```{r}
pred_probs_svm <-predict(svm_credit, credit_data, type = "prob")
svmLinearROC = roc(predictor=pred_probs_svm$yes,
                   response=credit_data$default,
                   levels=levels(credit_data$default),
                   direction = "<")
```
Decision tree
```{r}
pred_probs_tree <- predict(C5.0_credit, credit_data, type = "prob")
treeROC <- roc(credit_data$default, pred_probs_tree[, "yes"],
                levels = c('no','yes'),
                direction = '<')
```
ROC plots
```{r}
plot(treeROC, col="red")
plot(svmLinearROC, add=TRUE, col="blue")
plot(knnROC, add=TRUE, col="green")
```
Step 3

KNN auc
```{r}
auc(knnROC)
```
SVM linear auc
```{r}
auc(svmLinearROC)
```
Decision tree auc
```{r}
auc(treeROC)
```
We would choose the decision tree algorithm because it has the highest AUC (.865)

Step 4
```{r}
new_customers_data <- read.csv("new_customers.csv", stringsAsFactors = TRUE)
new_customers_data$credit_history <- ordered(new_customers_data$credit_history,
                                      levels = c("critical", "poor", "good","very good","perfect"))
new_customers_data$employment_duration <- 
  ordered(new_customers_data$employment_duration,
  levels = c("unemployed" , "< 1 year", "1 - 4 years", "4 - 7 years","> 7 years"))
```
KNN
```{r}
pred_knn_new_customers <-predict(knn_credit, newdata = new_customers_data)
pred_knn_new_customers
```
SVM linear
```{r}
pred_svm_new_customers <-predict(svm_credit, newdata = new_customers_data)
pred_svm_new_customers
```
Decision tree
```{r}
pred_decision_tree_new_customers <-predict(C5.0_credit, newdata = new_customers_data)
pred_decision_tree_new_customers
```
According to majority rule, the following prediction would be made for the five
customers in the new customers dataset: yes no yes no no (in order of customers listed
in the new customers dataset)

Step 5

Calculating individual accuracies via confusion matrix

KNN
```{r}
confusionMatrix(data = pred_knn_new_customers, new_customers_data$default)
```
SVM linear
```{r}
confusionMatrix(data = pred_svm_new_customers, new_customers_data$default)
```
Decision Tree
```{r}
confusionMatrix(data = pred_decision_tree_new_customers, new_customers_data$default)
```
Calculating accuracy for majority rule manually

Majority value predictions in order listed (predictions in brackets are correct):
[yes] [no] [yes] [no] [no]
Majority rule accuracy = 5/5 = 100%
The majority rule accuracy of 100% is greater than the individual accuracy of the algorithms 
which are all equal to 80%

